---
title: "analysis of growth form and leaf traits"
author: "Suz"
date: "22 July 2022"
output: html_document
---

Libraries required

```{r}
library(dplyr)
```

Read in relevant data

```{r}
growthformdata <- read.csv("growthform.csv")
leaftraitdata <- read.csv("leafdatafull.csv")
```

Stitch these two dataframes together and clean so that we have a dataframe that is just the log response ratios of each leaf trait for each species and their growth form. Also mutate a column for the mean trait change and the absolute mean trait change for the linear regression of this variable against growth form

```{r}
traitchangeandgrowthformdata <- left_join(growthformdata, leaftraitdata, by = "Species") %>%
  select(Species, growthform, thickLRR, lmaLRR, marginLRR, dimensionsLRR, stomataLRR, areaLRR, photoLRR, nitrogenLRR, WUELRR) %>%
  mutate(averagetraitchange = rowMeans(.[,c("thickLRR", "lmaLRR", "marginLRR", "dimensionsLRR", "stomataLRR", "areaLRR", "photoLRR", "nitrogenLRR", "WUELRR")], na.rm=T)) %>% #making an average trait change
  mutate(absoluteaveragetraitchange = abs(averagetraitchange)) %>% #and making an ABSOLUTE average trait change
  filter(!is.na(averagetraitchange)) #removing the species that had no data at all in this study - they may have been germinated and used in previous studies (e.g. Everingham et al 2021, Time travelling seeds reveal that plant regeneration and growth traits are responding to climate change, ECOLOGY, 2021)
```


Now running an analyis of growth form against average leaf trait change

```{r}
growthformlm <- lm(log(absoluteaveragetraitchange) ~ growthform, data = traitchangeandgrowthformdata) #absolute trait change is logged because of non-normally distributed data
summary(growthformlm)
```

Creating a map of our species with growth form and absolute change mapped onto these points - this was asked by reviewer for Ecology and Evolution


First constructing australian map as the background canvas of the map

```{r}
world <- ne_countries(scale = "medium", returnclass = "sf")

ausplot <- ggplot(data = world) +
  geom_sf(color = "white", size = 0.1, fill = "#bcc99c") +
  theme_classic()+
  theme(axis.line = element_line(colour = "white")) +
  coord_sf(xlim = c(110, 155), ylim = c(-45, -10.5), expand = FALSE)
ausplot

#save australia plot to create figure
ggsave(
  "australiamap.tiff",
  plot=ausplot,
  device = "tiff",
  width=10,
  height=8,
  dpi=200
)

ausplotzoomed <- ggplot(data = world) +
  geom_sf(color = "white", size = 0.1, fill = "#bcc99c") +
  theme_classic()+
  theme(axis.line = element_line(colour = "white")) +
  coord_sf(xlim = c(131, 155), ylim = c(-45, -25), expand = FALSE)
ausplotzoomed
```

### Then constructing Australian map with sites

```{r, warning=FALSE}
LeafChapterMapwithspecies <- ausplotzoomed +
  geom_point(data = traitchangeandgrowthformdata, aes(x= `Lon`, y = `Lat`, color = growthform, shape = growthform, size = (absoluteaveragetraitchange)), alpha = 0.8) +
  scale_size_continuous(range = c(2, 10)) +
  scale_shape_manual(values =c(15,16,17,18)) +
  scale_color_manual(values=c("darkseagreen4", "lightsalmon3", "deepskyblue4", "firebrick4"))

LeafChapterMapwithspecies

#save plot 
ggsave(
  "leafchaptermap.tiff",
  plot=LeafChapterMapwithspecies,
  device = "tiff",
  width=10,
  height=8,
  dpi=200
)
```

The editor of ecology and evolution also asked if our growth form analysis can consider generation time to determine if generation time has an impact on the amount of change a species can have- I believe there is limited data on generation time specifically but have a dataset from TRY of generation time

```{r}
generationtimeTRY <- read.delim("30063.txt") %>%
  rename(Species = SpeciesName)

generationtimeourspecies <- left_join(traitchangeandgrowthformdata, generationtimeTRY, by = "Species") %>%
  select(Species, OrigValueStr)
```

From looking at this dataset we can see we only have data on nine species: Bothriochloa macra, Senecio garlandii, Vittadinia cuneata (all herbs with <1 year to repro maturity), Allocasuarina nana, Melaleuca quinquenervia, Eucalyptus obliqua, Acacia baileyana, Allocasuarina littoralis and Acacia dealbata
<br>
Likewise, this data is not so accurate with many "X-Xyears" or "<Xyears" in these nine species - not that clear to use


Let's also try Austraits

```{r}
# load austraits
austraits <- load_austraits(version = "5.0.0", path = "data/austraits", update = TRUE)

# find traits for generation time or maturity or life span
lookup_trait(austraits, "matur") %>% head() #swap "matur" with lifespan or repro or otherwise

# from a quick look we can use "reproductive_maturity" and "lifespan"
# pull these traits from aus traits
data_repro_matur <- extract_trait(austraits, "reproductive_maturity") 
data_lifespan <- extract_trait(austraits, "lifespan")

# get traits for just our species 
data_repro_matur <- data_repro_matur$traits %>%
  rename(Species = taxon_name)
data_lifespan <- data_lifespan$traits %>%
  rename(Species = taxon_name)

#
austraits_our_data <- left_join(traitchangeandgrowthformdata, data_repro_matur, by = "Species") %>%
  left_join(., data_lifespan, by = "Species") %>%
  group_by(Species) %>%
  summarise(mean_repro_maturity = mean(value.x, na.rm = T),
            mean_lifespan = mean(value.y, na.rm = T)) %>%
  mutate(Reference = "AusTraits")

#so there are some more data from austraits but only 13 species
```

Now we combine what data we have from austraits, supplemented with any extra trait measures from TRY and run the same analyses to see if it impacts the amount of trait change

```{r}
repro_maturity_full <- left_join(traitchangeandgrowthformdata, austraits_our_data, by = "Species") %>%
  mutate(mean_repro_maturity = ifelse(Species == "Vittadinia cuneata", 1, mean_repro_maturity)) %>%
  mutate(Reference = ifelse(Species == "Vittadinia cuneata", "TRY", Reference)) #only extra measure from try is this one vittadinia species
```

Now run regression
```{r}
repro_maturity_trait_change <- lm(log(absoluteaveragetraitchange) ~ log(mean_repro_maturity), data = repro_maturity_full)
summary(repro_maturity_trait_change)
#still not significant - but not many measurements - and relationship (R squared) not strong either - no pattern
```

