---
title: "Analyses for two way regressions between leaf morphology and mean temp/precip"
author: "Suz Everingham"
date: "8 April 2020"
output: 
  html_document:
    keep_md: yes
    fig_crop: no
    highlight: zenburn
    theme: flatly
  
---

Libraries required

```{r, warning = FALSE, message=FALSE}
library(metafor)
library(ggplot2)
library(gridExtra)
```


Read in main data file - created in previous code: "reshaping data for leaf analysis"
```{r}
leafdata <- read.csv("leafdatafull.csv")
```

Paired analyses:

1. Leaf area vs mean temp // and Leaf area vs mean precip

```{r, warning = FALSE, message=FALSE}
leafareatemp <- rma(yi=areaLRR, vi = areavi, mods = Changeavtemp, data = leafdata)
```

```{r}
print(leafareatemp)
```

```{r, warning = FALSE, message=FALSE}
leafareaprecipt <- rma(yi=areaLRR, vi = areavi, mods = Changeav, data = leafdata)
```

```{r}
print(leafareaprecipt)
```

2. Leaf roundness (width:length) vs mean temp // and leaf roundness (width:length) vs mean precip

```{r, warning = FALSE, message=FALSE}
leafroundtemp <- rma(yi=dimensionsLRR, vi = roundnessvi, mods = Changeavtemp, data = leafdata) ###for some silly reason I interchanged the words "dimensions" and "roundness" when creating the dataframe previously - they mean the same thing
```
*Note that in the analysis I named the log response ratio "dimensions" and the variance variable "roundness" but these are interchangeable words for the same thing*
```{r}
print(leafroundtemp)
```

```{r, warning = FALSE, message=FALSE}
leafroundprecip <- rma(yi=dimensionsLRR, vi = roundnessvi, mods = Changeav, data = leafdata)
```

```{r}
print(leafroundprecip)
```

3. Leaf margin complexity (area:perimter) vs mean temp // and leaf margin complexity (area:perimeter) vs mean precip

```{r, warning = FALSE, message=FALSE}
leafmargintemp <- rma(yi=marginLRR, vi = marginvi, mods = Changeavtemp, data = leafdata)
```

```{r}
print(leafmargintemp)
```

```{r, warning = FALSE, message=FALSE}
leafmarginprecip <- rma(yi=marginLRR, vi = marginvi, mods = Changeav, data = leafdata)
```

```{r}
print(leafmarginprecip)
```

4. Leaf thickness vs mean temp // and leaf thickness vs mean precip

```{r, warning = FALSE, message=FALSE}
leafthicknesstemp <- rma(yi=thickLRR, vi = thicknessvi, mods = Changeavtemp, data = leafdata)
```

```{r}
print(leafthicknesstemp)
```

```{r, warning = FALSE, message=FALSE}
leafthicknessprecip <- rma(yi=thickLRR, vi = thicknessvi, mods = Changeav, data = leafdata)
```

```{r}
print(leafthicknessprecip)
```

Plotting relationships for figure 1
- Each relationship is plotted one by one between mean temperature and the leaf trait change metric and mean precipitation and the leaf trait change metric
- Colour scheme for manuscript is lightsalmon3 and darkseagreen4
- combine all plots together
- save tiff


1. Plots for both relationships for leaf surface area

```{r,  warning = FALSE, out.width=c('50%', '50%'), fig.show='hold'}
sizearea <- 1/sqrt(leafdata$areavi) ##to get the sizes for the points - based off variance from metafor for each within species change in trait

leafareatempplot <- ggplot(leafdata, aes(Changeavtemp, areaLRR)) + 
  geom_point(colour="darkseagreen4", size=sizearea*0.5) + 
  geom_abline(slope=-0.1073, intercept = -0.0620, colour="lightsalmon3", size = 2) +
  geom_vline(xintercept = 0, linetype = "dashed", alpha = 0.3, size = 1) +
  geom_hline(yintercept = 0, linetype = "dashed", alpha = 0.3, size= 1) +
  labs(cex=15, x="", y="Change in leaf area (cm²)") +
  theme_bw() +
  scale_y_continuous(labels = scales::number_format(accuracy = 0.1)) +
  theme(plot.background = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank() )+
  theme(panel.border= element_blank())+
  theme(axis.line = element_line(color="black", size = 0.5)) +
  theme(axis.text.y = element_text(size=18)) +
  theme(axis.text.x = element_blank())
plot(leafareatempplot)

leafareaprecipplot <- ggplot(leafdata, aes(Changeav, areaLRR)) + ##Changeav is change in average precipitation
  geom_point(colour="darkseagreen4", size=sizearea*0.5) +
  geom_vline(xintercept = 0, linetype = "dashed", alpha = 0.3, size=1) +
  geom_hline(yintercept = 0, linetype = "dashed", alpha = 0.3, size=1) +
  labs(cex=8, x="", y="") +
  theme_bw() +
  scale_y_continuous(labels = scales::number_format(accuracy = 0.1)) +
  theme(plot.background = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank() )+
  theme(panel.border= element_blank())+
  theme(axis.line = element_line(color="black", size = 0.5)) +
  theme(axis.text = element_blank())
plot(leafareaprecipplot)
```

2. Plots for both relationships for leaf roundness

```{r,  warning = FALSE, out.width=c('50%', '50%'), fig.show='hold'}
sizeroundness <- 1/sqrt(leafdata$roundnessvi)

roundnesstempplot <- ggplot(leafdata, aes(Changeavtemp, dimensionsLRR)) + 
  geom_point(colour="darkseagreen4", size=sizeroundness*0.5) +
  geom_vline(xintercept = 0, linetype = "dashed", alpha = 0.3, size=1) +
  geom_hline(yintercept = 0, linetype = "dashed", alpha = 0.3, size=1) +
  labs(cex=15, x="", y= expression(paste("Change in leaf roundness ", italic(" width:length"), " (cm)"))) +
  theme_bw() +
  scale_y_continuous(labels = scales::number_format(accuracy = 0.1)) +
  theme(plot.background = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank() )+
  theme(panel.border= element_blank())+
  theme(axis.line = element_line(color="black", size = 0.5)) +
  theme(axis.text.y = element_text(size=18)) +
  theme(axis.text.x = element_blank())
plot(roundnesstempplot)

roundnessprecipplot <- ggplot(leafdata, aes(Changeav, dimensionsLRR)) + 
  geom_point(colour="darkseagreen4", size=sizeroundness*0.5) +
  geom_vline(xintercept = 0, linetype = "dashed", alpha = 0.3, size=1) +
  geom_hline(yintercept = 0, linetype = "dashed", alpha = 0.3, size=1) +
  labs(cex=8, x="", y= "") +
  theme_bw() +
  scale_y_continuous(labels = scales::number_format(accuracy = 0.1)) +
  theme(plot.background = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank() )+
  theme(panel.border= element_blank())+
  theme(axis.line = element_line(color="black", size = 0.5)) +
  theme(axis.text = element_blank())
plot(roundnessprecipplot)

```

3. Plots for both relationships for leaf margin complexity

```{r, warning = FALSE, out.width=c('50%', '50%'), fig.show='hold'}
sizemargin <- 1/sqrt(leafdata$marginvi)

margintempplot <- ggplot(leafdata, aes(Changeavtemp, marginLRR)) + 
  geom_point(colour="darkseagreen4", size=sizemargin*0.5) +
  geom_abline(slope = 0.0472, intercept = 0.0276, colour="lightsalmon3", size = 2) +
  geom_vline(xintercept = 0, linetype = "dashed", alpha = 0.3, size=1) +
  geom_hline(yintercept = 0, linetype = "dashed", alpha = 0.3, size=1) +
  labs(cex=15, x="", y= expression(paste("Change in leaf margin complexity\n        [width:length (cm²/cm)]"))) +
  theme_bw() +
  scale_y_continuous(labels = scales::number_format(accuracy = 0.1)) +
  theme(plot.background = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank() )+
  theme(panel.border= element_blank())+
  theme(axis.line = element_line(color="black", size = 0.5)) +
  theme(axis.text.y = element_text(size=18)) +
  theme(axis.text.x = element_blank())
plot(margintempplot)

marginprecipplot <- ggplot(leafdata, aes(Changeav, marginLRR)) + 
  geom_point(colour="darkseagreen4", size=sizemargin*0.5) +
  geom_vline(xintercept = 0, linetype = "dashed", alpha = 0.3, size=1) +
  geom_hline(yintercept = 0, linetype = "dashed", alpha = 0.3, size=1) +
  labs(cex=8, x="", y= "") +
  theme_bw() +
  scale_y_continuous(labels = scales::number_format(accuracy = 0.1)) +
  theme(plot.background = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank() )+
  theme(panel.border= element_blank())+
  theme(axis.line = element_line(color="black", size = 0.5)) +
  theme(axis.text = element_blank())
plot(marginprecipplot)
```

4. Plots for both relationships of leaf thickness (NB the x axis labels finally appear here because these will be the bottom of the panel)

```{r, warning = FALSE, out.width=c('50%', '50%'), fig.show='hold'}
sizethickness <- 1/sqrt(leafdata$thicknessvi)
thicknesstempplot <- ggplot(leafdata, aes(Changeavtemp, thickLRR)) + 
  geom_point(colour="darkseagreen4", size=sizethickness*0.5) +
  geom_vline(xintercept = 0, linetype = "dashed", alpha = 0.3, size=1) +
  geom_hline(yintercept = 0, linetype = "dashed", alpha = 0.3, size=1) +
  labs(cex=20, x="Change in mean temperature (°C)", y= "Change in leaf thickness (mm)") +
  theme_bw() +
  scale_y_continuous(labels = scales::number_format(accuracy = 0.1)) +
  theme(plot.background = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank() )+
  theme(panel.border= element_blank())+
  theme(axis.line = element_line(color="black", size = 0.5)) +
  theme(axis.text = element_text(size=18)) +
  theme(axis.title.x = element_text(size=26))
plot(thicknesstempplot)
thicknessprecipplot <- ggplot(leafdata, aes(Changeav, thickLRR)) + 
  geom_point(colour="darkseagreen4", size=sizethickness*0.5) +
  geom_vline(xintercept = 0, linetype = "dashed", alpha = 0.3, size=1) +
  geom_hline(yintercept = 0, linetype = "dashed", alpha = 0.3, size=1) +
  labs(cex=20, x="Change in mean precipitation (mm)", y= "") +
  theme_bw() +
  scale_y_continuous(labels = scales::number_format(accuracy = 0.1)) +
  theme(plot.background = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank() )+
  theme(panel.border= element_blank())+
  theme(axis.line = element_line(color="black", size = 0.5)) +
  theme(axis.text.x = element_text(size=18)) +
  theme(axis.title.x = element_text(size=26)) +
  theme(axis.text.y = element_blank())
plot(thicknessprecipplot)
```

Join all plots together and save as tiff 

```{r figure1, dev = "png", fig.width = 12, fig.height=18, warning = FALSE}
figure1 <- grid.arrange(leafareatempplot, leafareaprecipplot,
                        roundnesstempplot, roundnessprecipplot,
                        margintempplot, marginprecipplot,
                        thicknesstempplot, thicknessprecipplot, ncol=2)

ggsave(
  "figure1.tiff",
  plot=figure1,
  device = "tiff",
  width=12,
  height=16,
  dpi=200
)

```

