---
title: "Analyses for two way regressions between leaf photosynthesis trait changes and change in mean temp/precip"
author: "Suz Everingham"
date: "16 June 2020"
output: 
  html_document:
    keep_md: yes
    fig_crop: no
    highlight: zenburn
    theme: flatly
---

Libraries required

```{r, warning = FALSE, message = FALSE}
library(metafor)
library(ggplot2)
library(gridExtra)
```


Read in main data file -  created in previous code: "reshaping data for leaf analysis"
```{r}
data <- read.csv("leafdatafull.csv")
```

Paired analyses:

1. Photosynthetic rate vs mean temp and photosynthetic rate vs mean precip

```{r, warning = FALSE}
phototemp <- rma(yi=photoLRR.y, vi = photosynthesisvi, mods = Changeavtemp, data = data)
```

```{r}
print(phototemp)
```

```{r, warning=FALSE}
photoprecipt <- rma(yi=photoLRR.y, vi = photosynthesisvi, mods = Changeav, data = data)
```

```{r}
print(photoprecipt)
```

2. stomata vs mean temp and stomata vs mean precip

```{r, warning=FALSE}
stomatatemp <- rma(yi=stomataLRR, vi = stomatavi, mods = Changeavtemp, data = data)
```

```{r}
print(stomatatemp)
```

```{r, warning=FALSE}
stomataprecipt <- rma(yi=stomataLRR, vi = stomatavi, mods = Changeav, data = data)
```

```{r}
print(stomataprecipt)
```

3. water use efficiency vs mean temp and water use efficiency vs mean precip

```{r, warning=FALSE}
WUEtemp <- rma(yi=WUELRR, vi = WUEvi, mods = Changeavtemp, data = data)
```

```{r}
print(WUEtemp)
```

```{r, warning=FALSE}
WUEprecipt <- rma(yi=WUELRR, vi = WUEvi, mods = Changeav, data = data)
```

```{r}
print(WUEprecipt)
```

Plotting relationships for figure 1
- Plot each relationship individually of change in photosynthetic trait by change in mean temp and then again by change in mean precipitation
- Colour scheme for thesis is lightsalmon3 and darkseagreen4
- combine all plots together
- save tiff


1. Plots for both relationships for photosynthesis

```{r, warning=FALSE, out.width=c('50%', '50%'), fig.show='hold'}
sizephoto <- 1/sqrt(data$photosynthesisvi) ##to get the sizes for the points - based off variance from metafor for each within species change in trait

phototempplot <- ggplot(data=data, aes(Changeavtemp, photoLRR.x)) + 
  geom_point(colour="darkseagreen4", size=sizephoto*0.5) +
    geom_vline(xintercept = 0, linetype = "dashed", alpha = 0.3, size=1) +
  geom_hline(yintercept = 0, linetype = "dashed", alpha = 0.3, size=1) +
  labs(cex=15, x="", y="Change in photosynthetic rate ()") +
  theme_bw() +
  theme(plot.background = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank() )+
  theme(panel.border= element_blank())+
  theme(axis.line = element_line(color="black", size = 0.5)) +
  theme(axis.text.y = element_text(size=18)) +
  theme(axis.text.x = element_blank())
plot(phototempplot)

photoprecipplot <- ggplot(data=data, aes(Changeav, photoLRR.x)) + 
  geom_point(colour="darkseagreen4", size=sizephoto*0.5) +
  geom_vline(xintercept = 0, linetype = "dashed", alpha = 0.3, size=1) +
  geom_hline(yintercept = 0, linetype = "dashed", alpha = 0.3, size=1) +
  labs(cex=8, x="", y="") +
  theme_bw() +
  theme(plot.background = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank() )+
  theme(panel.border= element_blank())+
  theme(axis.line = element_line(color="black", size = 0.5)) +
  theme(axis.text = element_blank())+
  coord_cartesian(xlim=c(-0.35, 0.2))
plot(photoprecipplot)
```

2. water use efficiency and mean temp and mean precip

```{r, warning=FALSE, out.width=c('50%', '50%'), fig.show='hold'}
sizeWUE <- 1/sqrt(data$WUEvi) ##to get the sizes for the points - based off variance from metafor for each within species change in trait

WUEtempplot <- ggplot(data=data, aes(Changeavtemp, WUELRR)) + 
  geom_point(colour="darkseagreen4", size=sizeWUE*0.5) +
  labs(cex=15, x="", y="Change in water use efficiency (Âµmol COâ mmolâ»Â¹ HâO)") +
  geom_vline(xintercept = 0, linetype = "dashed", alpha = 0.3, size=1) +
  geom_hline(yintercept = 0, linetype = "dashed", alpha = 0.3, size=1) +
  theme_bw() +
  theme(plot.background = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank() )+
  theme(panel.border= element_blank())+
  theme(axis.line = element_line(color="black", size = 0.5)) +
  theme(axis.text.y = element_text(size=18)) +
  theme(axis.text.x = element_blank()) +
  scale_y_continuous(labels = scales::number_format(accuracy = 0.1))
plot(WUEtempplot)

WUEprecipplot <- ggplot(data=data, aes(Changeav, WUELRR)) + 
  geom_point(colour="darkseagreen4", size=sizeWUE*0.5) +
  geom_vline(xintercept = 0, linetype = "dashed", alpha = 0.3, size=1) +
  geom_hline(yintercept = 0, linetype = "dashed", alpha = 0.3, size=1) +
  labs(cex=8, x="", y="") +
  theme_bw() +
  theme(plot.background = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank() )+
  theme(panel.border= element_blank())+
  theme(axis.line = element_line(color="black", size = 0.5)) +
  theme(axis.text = element_blank()) +
  coord_cartesian(xlim=c(-0.35, 0.2)) +
  scale_y_continuous(labels = scales::number_format(accuracy = 0.1))
plot(WUEprecipplot)
```

3. for stomatal density and mean temp and mean precip

```{r, warning=FALSE, out.width=c('50%', '50%'), fig.show='hold'}
sizestomata <- 1/sqrt(data$stomatavi) ##to get the sizes for the points - based off variance from metafor for each within species change in trait

stomatatempplot <- ggplot(data=data, aes(Changeavtemp, stomataLRR)) + 
  labs(cex=15, x="Change in mean temperature (Â°C)", y="Change in stomatal density (mmÂ²)") +
  geom_vline(xintercept = 0, linetype = "dashed", alpha = 0.3, size=1) +
  geom_hline(yintercept = 0, linetype = "dashed", alpha = 0.3, size=1) +
  geom_point(colour="darkseagreen4", size=sizestomata*0.5) +
  theme_bw() +
  theme(plot.background = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank() )+
  theme(panel.border= element_blank())+
  theme(axis.line = element_line(color="black", size = 0.5)) +
  theme(axis.text.y = element_text(size=18)) +
  theme(axis.text.x = element_text(size=18),
        axis.title.x = element_text(size=20))
plot(stomatatempplot)

stomataprecipplot <- ggplot(data=data, aes(Changeav, stomataLRR)) + 
  geom_point(colour="darkseagreen4", size=sizestomata*0.5) +
  geom_vline(xintercept = 0, linetype = "dashed", alpha = 0.3, size=1) +
  geom_hline(yintercept = 0, linetype = "dashed", alpha = 0.3, size=1) +
  labs(x="Change in mean precipitation (mm)", y="") +
  theme_bw() +
  theme(plot.background = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank() )+
  theme(panel.border= element_blank())+
  theme(axis.line = element_line(color="black", size = 0.5)) +
  theme(axis.text.x = element_text(size=18),
        axis.text.y = element_blank(),
        axis.title.x = element_text(size =20)) +
  coord_cartesian(xlim=c(-0.35, 0.2))
plot(stomataprecipplot)
```

Stitch all the plots together and save as a tiff file

```{r, warning=FALSE}
figure2 <- grid.arrange(phototempplot, photoprecipplot,
                        WUEtempplot, WUEprecipplot,
                        stomatatempplot, stomataprecipplot, ncol=2)

ggsave(
  "figure2.tiff",
  plot=figure2,
  device = "tiff",
  width=12,
  height=13,
  dpi=200
)
```

