---
title: "meta-analyses of leaf traits and climate variables"
author: "Suz"
date: "23 July 2020"
output: html_document
---

libraries required

```{r}
library(metafor)
library(glmulti)
library(ggplot2)
library(gridExtra)
```

read in data

```{r}
leafdata <- read.csv("leafdatafull.csv")
```

Create create function for rma.glmulti- rma with yi and vi

```{r}
rma.glmulti <- function(formula, vi, data, ...) {
  rma(as.formula(paste(deparse(formula))), vi=vi, data=data, method = "ML", ...)
}
```

Also create a function for getting the moderator importance and significance averaged across all models
```{r}
setMethod('getfit', 'rma.uni', function(object, ...){
  if (object$test == "z") {
    cbind(estimate = coef(object), se = sqrt(diag(vcov(object))), df= 10000)
  } else {
    cbind(estimate = coef(object), se = sqrt(diag(vcov(object))), df= object$k - object$p)
  }
})
```


Run meta-analytic multi-model selection on each leaf trait and leaf ecnomics against all climate variables

1. Leaf area
```{r}
leafareamodelselection <- glmulti(areaLRR ~ Changeavtemp + Changetemprange + Changetempvar + Changeav + Changerange + Changevar + Changemaxseasonal + Changeminseasonal + Changedrought + Changeheatwave + Changevpd + Changemaxdryspell + Changedrought, vi = leafdata$areavi, data=leafdata, level=1, fitfunction=rma.glmulti, crit="aicc", confsetsize = 8192)

summary(leafareamodelselection@objects[[1]])
print(leafareamodelselection)
plot(leafareamodelselection)

mmi <- as.data.frame(coef(leafareamodelselection))
mmi <- data.frame(Estimate=mmi$Est, SE=sqrt(mmi$Uncond), Importance=mmi$Importance, row.names=row.names(mmi))
mmi$z <- mmi$Estimate / mmi$SE
mmi$p <- 2*pnorm(abs(mmi$z), lower.tail=FALSE)
names(mmi) <- c("Estimate", "Std. Error", "Importance", "z value", "Pr(>|z|)")
mmi$ci.lb <- mmi[[1]] - qnorm(.975) * mmi[[2]]
mmi$ci.ub <- mmi[[1]] + qnorm(.975) * mmi[[2]]
mmi <- mmi[order(mmi$Importance, decreasing=TRUE), c(1,2,4:7,3)]
round(mmi, 4)
```

2. Leaf roundness

```{r}
leafroundnessmodelselection <- glmulti(dimensionsLRR ~ Changeavtemp + Changetemprange + Changetempvar + Changeav + Changerange + Changevar + Changemaxseasonal + Changeminseasonal + Changedrought + Changeheatwave + Changevpd + Changemaxdryspell + Changedrought, vi = leafdata$roundnessvi, data=leafdata, level=1, fitfunction=rma.glmulti, crit="aicc", confsetsize = 8192)

summary(leafroundnessmodelselection@objects[[1]])
print(leafroundnessmodelselection)
plot(leafroundnessmodelselection)

mmi <- as.data.frame(coef(leafroundnessmodelselection))
mmi <- data.frame(Estimate=mmi$Est, SE=sqrt(mmi$Uncond), Importance=mmi$Importance, row.names=row.names(mmi))
mmi$z <- mmi$Estimate / mmi$SE
mmi$p <- 2*pnorm(abs(mmi$z), lower.tail=FALSE)
names(mmi) <- c("Estimate", "Std. Error", "Importance", "z value", "Pr(>|z|)")
mmi$ci.lb <- mmi[[1]] - qnorm(.975) * mmi[[2]]
mmi$ci.ub <- mmi[[1]] + qnorm(.975) * mmi[[2]]
mmi <- mmi[order(mmi$Importance, decreasing=TRUE), c(1,2,4:7,3)]
round(mmi, 4)
```

3. Margin complexity

```{r}
leafmarginmodelselection <- glmulti(marginLRR ~ Changeavtemp + Changetemprange + Changetempvar + Changeav + Changerange + Changevar + Changemaxseasonal + Changeminseasonal + Changedrought + Changeheatwave + Changevpd + Changemaxdryspell + Changedrought, vi = leafdata$marginvi, data=leafdata, level=1, fitfunction=rma.glmulti, crit="aicc", confsetsize = 8192)

summary(leafmarginmodelselection@objects[[1]])
print(leafmarginmodelselection)
plot(leafmarginmodelselection)

mmi <- as.data.frame(coef(leafmarginmodelselection))
mmi <- data.frame(Estimate=mmi$Est, SE=sqrt(mmi$Uncond), Importance=mmi$Importance, row.names=row.names(mmi))
mmi$z <- mmi$Estimate / mmi$SE
mmi$p <- 2*pnorm(abs(mmi$z), lower.tail=FALSE)
names(mmi) <- c("Estimate", "Std. Error", "Importance", "z value", "Pr(>|z|)")
mmi$ci.lb <- mmi[[1]] - qnorm(.975) * mmi[[2]]
mmi$ci.ub <- mmi[[1]] + qnorm(.975) * mmi[[2]]
mmi <- mmi[order(mmi$Importance, decreasing=TRUE), c(1,2,4:7,3)]
round(mmi, 4)
```

4. Leaf thickness

```{r}
leafthicknessmodelselection <- glmulti(thickLRR ~ Changeavtemp + Changetemprange + Changetempvar + Changeav + Changerange + Changevar + Changemaxseasonal + Changeminseasonal + Changedrought + Changeheatwave + Changevpd + Changemaxdryspell + Changedrought, vi = leafdata$thicknessvi, data=leafdata, level=1, fitfunction=rma.glmulti, crit="aicc", confsetsize = 8192)

summary(leafthicknessmodelselection@objects[[1]])
print(leafthicknessmodelselection)
plot(leafthicknessmodelselection)

mmi <- as.data.frame(coef(leafthicknessmodelselection))
mmi <- data.frame(Estimate=mmi$Est, SE=sqrt(mmi$Uncond), Importance=mmi$Importance, row.names=row.names(mmi))
mmi$z <- mmi$Estimate / mmi$SE
mmi$p <- 2*pnorm(abs(mmi$z), lower.tail=FALSE)
names(mmi) <- c("Estimate", "Std. Error", "Importance", "z value", "Pr(>|z|)")
mmi$ci.lb <- mmi[[1]] - qnorm(.975) * mmi[[2]]
mmi$ci.ub <- mmi[[1]] + qnorm(.975) * mmi[[2]]
mmi <- mmi[order(mmi$Importance, decreasing=TRUE), c(1,2,4:7,3)]
round(mmi, 4)
```

5. Photosynthetic rate

```{r}
photosynthesismodelselection <- glmulti(photoLRR.x ~ Changeavtemp + Changetemprange + Changetempvar + Changeav + Changerange + Changevar + Changemaxseasonal + Changeminseasonal + Changedrought + Changeheatwave + Changevpd + Changemaxdryspell + Changedrought, vi = leafdata$photosynthesisvi, data=leafdata, level=1, fitfunction=rma.glmulti, crit="aicc", confsetsize = 8192)

summary(photosynthesismodelselection@objects[[1]])
print(photosynthesismodelselection)
plot(photosynthesismodelselection)

mmi <- as.data.frame(coef(photosynthesismodelselection))
mmi <- data.frame(Estimate=mmi$Est, SE=sqrt(mmi$Uncond), Importance=mmi$Importance, row.names=row.names(mmi))
mmi$z <- mmi$Estimate / mmi$SE
mmi$p <- 2*pnorm(abs(mmi$z), lower.tail=FALSE)
names(mmi) <- c("Estimate", "Std. Error", "Importance", "z value", "Pr(>|z|)")
mmi$ci.lb <- mmi[[1]] - qnorm(.975) * mmi[[2]]
mmi$ci.ub <- mmi[[1]] + qnorm(.975) * mmi[[2]]
mmi <- mmi[order(mmi$Importance, decreasing=TRUE), c(1,2,4:7,3)]
round(mmi, 4)
```

6. Water use efficiency

```{r}
WUEmodelselection <- glmulti(WUELRR ~ Changeavtemp + Changetemprange + Changetempvar + Changeav + Changerange + Changevar + Changemaxseasonal + Changeminseasonal + Changedrought + Changeheatwave + Changevpd + Changemaxdryspell + Changedrought, vi = leafdata$WUEvi, data=leafdata, level=1, fitfunction=rma.glmulti, crit="aicc", confsetsize = 8192)

summary(WUEmodelselection@objects[[1]])
print(WUEmodelselection)
plot(WUEmodelselection)

mmi <- as.data.frame(coef(WUEmodelselection))
mmi <- data.frame(Estimate=mmi$Est, SE=sqrt(mmi$Uncond), Importance=mmi$Importance, row.names=row.names(mmi))
mmi$z <- mmi$Estimate / mmi$SE
mmi$p <- 2*pnorm(abs(mmi$z), lower.tail=FALSE)
names(mmi) <- c("Estimate", "Std. Error", "Importance", "z value", "Pr(>|z|)")
mmi$ci.lb <- mmi[[1]] - qnorm(.975) * mmi[[2]]
mmi$ci.ub <- mmi[[1]] + qnorm(.975) * mmi[[2]]
mmi <- mmi[order(mmi$Importance, decreasing=TRUE), c(1,2,4:7,3)]
round(mmi, 4)
```

7. Stomatal density

```{r}
stomatamodelselection <- glmulti(stomataLRR ~ Changeavtemp + Changetemprange + Changetempvar + Changeav + Changerange + Changevar + Changemaxseasonal + Changeminseasonal + Changedrought + Changeheatwave + Changevpd + Changemaxdryspell + Changedrought, vi = leafdata$stomatavi, data=leafdata, level=1, fitfunction=rma.glmulti, crit="aicc", confsetsize = 8192)

summary(stomatamodelselection@objects[[1]])
print(stomatamodelselection)
plot(stomatamodelselection)

mmi <- as.data.frame(coef(stomatamodelselection))
mmi <- data.frame(Estimate=mmi$Est, SE=sqrt(mmi$Uncond), Importance=mmi$Importance, row.names=row.names(mmi))
mmi$z <- mmi$Estimate / mmi$SE
mmi$p <- 2*pnorm(abs(mmi$z), lower.tail=FALSE)
names(mmi) <- c("Estimate", "Std. Error", "Importance", "z value", "Pr(>|z|)")
mmi$ci.lb <- mmi[[1]] - qnorm(.975) * mmi[[2]]
mmi$ci.ub <- mmi[[1]] + qnorm(.975) * mmi[[2]]
mmi <- mmi[order(mmi$Importance, decreasing=TRUE), c(1,2,4:7,3)]
round(mmi, 4)
```

8 (and final). leaf economics

```{r}
leafeconomicsmodelselection <- glmulti(PC1 ~ Changeavtemp + Changetemprange + Changetempvar + Changeav + Changerange + Changevar + Changemaxseasonal + Changeminseasonal + Changedrought + Changeheatwave + Changevpd + Changemaxdryspell + Changedrought, vi = 1, data=leafdata, level=1, fitfunction=rma.glmulti, crit="aicc", confsetsize = 8192)

summary(leafeconomicsmodelselection@objects[[1]])
print(leafeconomicsmodelselection)
plot(leafeconomicsmodelselection)

mmi <- as.data.frame(coef(leafeconomicsmodelselection))
mmi <- data.frame(Estimate=mmi$Est, SE=sqrt(mmi$Uncond), Importance=mmi$Importance, row.names=row.names(mmi))
mmi$z <- mmi$Estimate / mmi$SE
mmi$p <- 2*pnorm(abs(mmi$z), lower.tail=FALSE)
names(mmi) <- c("Estimate", "Std. Error", "Importance", "z value", "Pr(>|z|)")
mmi$ci.lb <- mmi[[1]] - qnorm(.975) * mmi[[2]]
mmi$ci.ub <- mmi[[1]] + qnorm(.975) * mmi[[2]]
mmi <- mmi[order(mmi$Importance, decreasing=TRUE), c(1,2,4:7,3)]
round(mmi, 4)

#and again with principal component 2

leafeconomicsmodelselection2 <- glmulti(PC2 ~ Changeavtemp + Changetemprange + Changetempvar + Changeav + Changerange + Changevar + Changemaxseasonal + Changeminseasonal + Changedrought + Changeheatwave + Changevpd + Changemaxdryspell + Changedrought, vi = 1, data=leafdata, level=1, fitfunction=rma.glmulti, crit="aicc", confsetsize = 8192)

summary(leafeconomicsmodelselection2@objects[[1]])
print(leafeconomicsmodelselection2)
plot(leafeconomicsmodelselection2)

mmi <- as.data.frame(coef(leafeconomicsmodelselection2))
mmi <- data.frame(Estimate=mmi$Est, SE=sqrt(mmi$Uncond), Importance=mmi$Importance, row.names=row.names(mmi))
mmi$z <- mmi$Estimate / mmi$SE
mmi$p <- 2*pnorm(abs(mmi$z), lower.tail=FALSE)
names(mmi) <- c("Estimate", "Std. Error", "Importance", "z value", "Pr(>|z|)")
mmi$ci.lb <- mmi[[1]] - qnorm(.975) * mmi[[2]]
mmi$ci.ub <- mmi[[1]] + qnorm(.975) * mmi[[2]]
mmi <- mmi[order(mmi$Importance, decreasing=TRUE), c(1,2,4:7,3)]
round(mmi, 4)

```

Plotting the two leaf margin relationships that came out significant

```{r}
leafmarginandtempvar <- rma(yi=marginLRR, vi = marginvi, mods = Changetempvar, data=leafdata)
print(leafmarginandtempvar)

#slope = 2.5664
#intercept = 0.0049

sizemargin <- 1/sqrt(leafdata$marginvi)

margintempvarplot <- ggplot(data=leafdata, aes(Changetempvar, y=marginLRR)) + 
  geom_point(colour="lightsalmon3", size=sizemargin*0.5) +
  labs(cex=15, x="", y= expression(paste("Change in leaf margin complexity\n[width:length (cm²/cm)]"))) +
   geom_abline(slope=2.5664, intercept=0.0049, colour="darkseagreen4", size = 2) +
  geom_vline(xintercept = 0, linetype = "dashed", alpha = 0.3, size=1) +
  geom_hline(yintercept = 0, linetype = "dashed", alpha = 0.3, size=1) +
  theme_bw() +
  theme(plot.background = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank() )+
  theme(panel.border= element_blank())+
  theme(axis.line = element_line(color="black", size = 0.5)) +
  theme(axis.text.y = element_text(size=18)) +
  theme(axis.text.x = element_text(size=18))
plot(margintempvarplot)

## and now leaf margin nad dry spell duration
leafmarginanddryspell <- rma(yi=marginLRR, vi = marginvi, mods = Changemaxdryspell, data=leafdata)
print(leafmarginanddryspell)

#slope = -0.0490
#intercept = 0.0209

margindryspellplot <- ggplot(leafdata, aes(Changemaxdryspell, marginLRR)) + 
  geom_point(colour="lightsalmon3", size=sizemargin*0.5) +
  labs(cex=15, x="Dry spell", y= "") +
   geom_abline(slope=-0.0490, intercept=0.0209, colour="darkseagreen4", size = 2) +
  geom_vline(xintercept = 0, linetype = "dashed", alpha = 0.3, size=1) +
  geom_hline(yintercept = 0, linetype = "dashed", alpha = 0.3, size=1) +
  theme_bw() +
  theme(plot.background = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank() )+
  theme(panel.border= element_blank())+
  theme(axis.line = element_line(color="black", size = 0.5)) +
  theme(axis.text.y = element_blank()) +
  theme(axis.text.x = element_text(size=18))
plot(margindryspellplot)

```

Join all the graphs together and save as tiff

```{r}
figure4 <- grid.arrange(margintempvarplot, margindryspellplot, ncol=2)

ggsave(
  "figure4.tiff",
  plot=figure4,
  device = "tiff",
  width=16,
  height=6,
  dpi=200
)
```

