---
title: "PCA analyses for leaf economics traits"
author: "Suz Everingham"
date: "22 June 2020"
output: 
  html_document:
    keep_md: yes
    fig_crop: no
    highlight: zenburn
    theme: flatly
---
libraries required

```{r}
library(missMDA)
library(FactoMineR)
library(dplyr)
library(tidyr)
library(paran)
```

Reading in Data created previously in rmarkdown 
```{r}
leafdata <- read.csv("leafdatafull.csv")
```

Create a dataframe with just the three variables to see where the NAs lie 

```{r}
leafeconomicdata <- leafdata[c("Species", "Changeavtemp", "Changetempvar", "Changetemprange", "Changeav", "Changevar", "Changerange", "Changedrought", "Changemaxseasonal", "Changeminseasonal", "Changeheatwave", "Changevpd", "Changemaxdryspell", "photoLRR.x", "photosynthesisvi", "lmainverseLRR.x", "lmainversevi", "nitrogenLRR.x","nitrogenvi")]
```

A few species have no data at all (due to lack of germination or death in glasshouse). These species we will remove as they are of no use in a PCA. There are 8 species that need some data imputation, which isn't too bad.

Create a wide data frame so that each species gets a PCA for the three variables

```{r}
#select the data of species and the three traits

leafeconomicPCAdata1 <- leafdata[c("Species", "photoLRR.x", "nitrogenLRR.x", "lmainverseLRR.x")]

```

Now I want to remove the species that have NAs across all three leaf economic traits - these cannot be imputed and do not contribute any data

```{r}

leafeconomicPCAdataspecies <- filter(leafeconomicPCAdata1, Species != "Acacia aneura" & Species != "Acronychia imperforata" & Species != "Allocasuarina monilifera" & Species != "Austrostipa densiflora" & Species != "Callitris glaucophylla" & Species != "Crowea exalata" & Species != "Eucalyptus fastigata" & Species != "Eucalyptus langleyi" & Species != "Eucalyptus pauciflora" & Species != "Gompholobium latifolium" & Species != "Grevillea mucronulata" & Species != "Polycalymma stuartii" & Species != "Xerochrysum viscosum")

leafeconomicPCAdatafinal <- leafeconomicPCAdataspecies %>% select(2:4) ## selecting only the numeric columns
```


Run a PCA analysis on the leaf economic trait variables - photosynthesis, LMA (inverse- as it decreases as leaf economic traits increase) and leaf nitrogen content

```{r}
leafeconomicPCAimputed <- imputePCA(leafeconomicPCAdatafinal)

leafeconomi.PCA <- prcomp(t(leafeconomicPCAimputed$completeObs))

plot(leafeconomi.PCA)
summary(leafeconomi.PCA)
str(leafeconomi.PCA)

```

Check that I should only use PC1 using a parallel analysis of the PCA

```{r}
paran(leafeconomicPCAimputed$completeObs, iterations = 1000)
```

Save that PCA analysis as a dataframe and bind with the species names and data 

```{r}
leafeconomics <- cbind(leafeconomicPCAdataspecies,leafeconomi.PCA$rotation)
```

save that data as a csv to use in further analysis

```{r}
write.csv(leafeconomics, "leafeconomics.csv")
```

